---
title: "PML - Prediction Assigment"
author: "SK"
date: "23 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Housekeeping
Preapre environment for analysis
```{r, message=F, warning=F}
library(caret)
library(randomForest)
set.seed(123)
getwd()
```

## Read the data
```{r, message=F, warning=F}
data.in.training <- read.csv("pml-training.csv", na.strings = c("NA","","#DIV/0!"), stringsAsFactors = FALSE)
data.in.testing <- read.csv("pml-testing.csv", na.strings = c("NA","","#DIV/0!"), stringsAsFactors = FALSE)
```

## Have a look at the data
```{r, message=F, warning=F}
dim(data.in.training)
str(data.in.training)
```

## Remove aggregated data
```{r, message=F, warning=F}
data.in <- data.in.training[!(data.in.training$new_window=="yes"),]
```
## Remove variables with a lot of NAs
```{r, message=F, warning=F}
counts.NA <- sapply(data.in.training, function(x) sum(is.na(x)))
to.remove <- (counts.NA/nrow(data.in.training))>0.2
to.remove[1] <- TRUE # Remove ID
data.in <- data.in.training[,!(to.remove)]
```

## Find arm - belt - dumbell - forearm variables
```{r, message=F, warning=F}
vars.all <- names(data.in)
vars.all
vars.arm <- vars.all[grepl("_arm", vars.all, fixed=TRUE)]
vars.belt <- vars.all[grepl("_belt", vars.all, fixed=TRUE)]
vars.dumbell <- vars.all[grepl("_dumbbell", vars.all, fixed=TRUE)]
vars.forearm <- vars.all[grepl("_forearm", vars.all, fixed=TRUE)]
```

## Create a dataset for analysis with accelerometers data + name + target variable
```{r, message=F, warning=F}
data.in <- data.in[c(vars.arm,vars.belt,vars.dumbell,vars.forearm,"user_name","classe")]
```

## Convert string variables to factors
## Find all string variables
```{r, message=F, warning=F}
is.string.var <- colnames(data.in)[sapply(data.in, is.character)]
data.in[is.string.var] <- lapply(data.in[is.string.var], as.factor)
data.in.testing$user_name <- as.factor(data.in.testing$user_name)
str(data.in)
```

## Split the file to Train - Test (stratified - by default).
## I want a least 1000 records for each class in the test set
```{r, message=F, warning=F}
table(data.in$classe)
perc.train <-1-1000/min(table(data.in$classe))
trainIndex = createDataPartition(data.in$classe, p = perc.train,list=FALSE)
df.train = data.in[trainIndex,]
df.test = data.in[-trainIndex,]
table(df.test$classe)
```

## Run cross-validation. Calculate The Best Tune of number of variable randomly sampled.
```{r, message=F, warning=F}
fitControl = trainControl( method = "cv", number = 5)
fit.RF.cv <- train(classe ~ ., data = df.train, method = "rf", trControl = fitControl)
fit.RF.cv
```
## Average accuracy is 0.9911
```{r, message=F, warning=F}
confusionMatrix.train(fit.RF.cv)
```

## Build random forest model
```{r, message=F, warning=F}
fit.RF <- randomForest(classe~.,data=df.train,  mtry = fit.RF.cv$bestTune$mtry)
fit.RF
```

## OOB estimate of  error rate is 0.68%. Check it on test data
```{r, message=F, warning=F}
confusionMatrix(predict(fit.RF,newdata=df.test), df.test$classe)$overall["Accuracy"]
```
## Results are close

### Let's predict tesitn data for submission
```{r, message=F, warning=F}
ForSubmission = predict(fit.RF, newdata = data.in.testing)
ForSubmission
```

## We got all 20 cases right!

## Let's check LDA and gbm
```{r, message=F, warning=F}
fit.LDA <- train(classe ~ ., data = df.train, method = "lda")
confusionMatrix(predict(fit.LDA,newdata=df.test), df.test$classe)$overall["Accuracy"]

fit.gbm <- train(classe ~ ., data = df.train, method = "gbm", verbose=FALSE)
confusionMatrix(predict(fit.gbm,newdata=df.test), df.test$classe)$overall["Accuracy"]
```
## RF outperforms LDA and GBM for this data set

## As a conclusion
According to Breiman and Cutler we don't need cross-validation for RF
<https://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr>

In random forests, there is no need for cross-validation or a separate test set to get an unbiased estimate 
of the test set error. It is estimated internally, during the run, as follows:
Each tree is constructed using a different bootstrap sample from the original data. About one-third of the 
cases are left out of the bootstrap sample and not used in the construction of the kth tree.
Put each case left out in the construction of the kth tree down the kth tree to get a classification. 
In this way, a test set classification is obtained for each case in about one-third of the trees. 
At the end of the run, take j to be the class that got most of the votes every time case n was oob. 
The proportion of times that j is not equal to the true class of n averaged over all cases is 
the oob error estimate. This has proven to be unbiased in many tests.
